---
title: "Hypothesis Testing and Confidence Intervals"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      bg: "#FFFFFF"
      fg: "#000000" 
      primary: "#099587"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(thematic)
thematic::thematic_rmd()
```

###  `cars` data

```{r cars, fig.cap ="Scatterplot of cars data"}
library(tidyverse)
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() + 
  theme_minimal() +
  labs(x="Speed", y="Stopping Distance")
```

*** 

I will work with R's internal dataset on cars: `cars`.  There are two variables in the dataset, speed [speed] and distance [dist] this is what they look like.

We require two things:

1. A **claim** based on **target/hypothesized value**
   + Alternative two-sided: $\mu = target$ vs. $\mu \neq target$
   + Alternative greater: $\mu \leq target$ vs. $\mu > target$
   + Alternative lesser: $\mu \geq target$ vs. $\mu < target$

2. A **confidence level** that is a threshold in probability for rendering a true/false answer.

### Hypothesis Testing and $t$

I will work with the speed variable.  Let me consider two distinct hypotheses though this is not entirely proper in the sense that any given application of hypothesis testing should have a clear hypothesis about the variable of interest rather than waffling amongst two as I will.  I do so only for illustrative purposes.  To begin, I need to specify both the hypothesis and the confidence level that I intend to use to evaluate it.  

- **Case 1**: The hypothesis to advance is that 17 is the true average speed.  The alternative must then be that the average speed is not 17; it could be smaller or larger than 17. I wish to evaluate this claim with 90% confidence.  
- **Case 2**: The hypothesis to advance is that 17 or greater is the true average speed.  The alternative must then be that the average speed is less than 17.  I wish to evaluate this claim with 90% confidence.  
- **Case 1**: The hypothesis to advance is that 17 or less is the true average speed.  The alternative must then be that the average speed is greater than 17.  I wish to evaluate this claim with 90% confidence.  


The $t$ equation is given by:

$$t=\frac{\overline{x} - \mu}{\frac{s}{\sqrt{n}}}$$

One further algebraic manipulation before starting.  Let's solve for $\overline{x}$.

$$\mu + t\left(\frac{s}{\sqrt{n}}\right)=\overline{x}$$

This fully defines the parts we require and the core problem

### Details of the Data {data-commentary-width=600}

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
cars %>% summarise(Mean = mean(speed),
                   SD = sd(speed),
                   SE = sd(speed)/sqrt(n()),
                   N = n())
```

***

- The mean of speed is 15.4 mph
- The standard deviation is 5.288 mph.
- There are fifty observations. [n=50]
- The standard error of the mean is 0.7478.

$$t = \frac{15.4 - 17}{\frac{s}{\sqrt{n}}} = \frac{-1.6}{0.7478}=-2.14$$

### Case 1: A Two-sided Alternative {data-commentary-width=400}

```{r, warning=FALSE, message=FALSE}
library(radiant)
result <- prob_tdist(df = 49, plb = 0.05, pub=0.95)
# summary(result, type = "probs")
p1 <- plot(result, type = "probs") + 
  theme_minimal() +
  geom_label(aes(x=-1.677, y=0.25, label="t=-1.677")) +
  geom_label(aes(x=1.677, y=0.25, label="t=1.677")) +
  geom_label(aes(x=0, y=0.15, label="p=0.9"), size=2) +
  geom_label(aes(x=-3, y=0.075, label="p=0.05"), size=2) +
  geom_label(aes(x=3, y=0.075, label="p=0.05"), size=2) +
  labs(title="Critical Values", x="t")  +
  xlim(-4,4)
result2 <- prob_tdist(df = 49, lb = -2.14, ub=2.14)
p2 <- plot(result2) + theme_minimal() +
  geom_vline(aes(xintercept=-2.14), color="#099587") +
  geom_vline(aes(xintercept=2.14), color="#099587")  +
  geom_label(aes(x=-2.14, y=0.25, label="t=-2.14")) +
  geom_label(aes(x=2.14, y=0.25, label="t=2.14")) +
  geom_label(aes(x=0, y=0.15, label="p=0.963"), size=2) +
  geom_label(aes(x=-3, y=0.075, label="p=0.0187"), size=2) +
  geom_label(aes(x=3, y=0.075, label="p=0.0187"), size=2) +
  guides(size="none") +
  labs(title="Result", x="t") + 
  xlim(-4,4)
library(patchwork)
p1 + p2
```

***

```
Critical Values
P(-1.677 < X < 1.677)     = 0.9
1 - P(-1.677 < X < 1.677) = 0.1

Result
P(X < -2.14) = 0.019
P(X > 2.14) = 0.019
P(-2.14 < X < 2.14)     = 0.963
1 - P(-2.14 < X < 2.14) = 0.037
```


***

Knowing only the sample size, 50, is sufficient to determine what $t$ must be to reject a mean of 17 in favor of the alternative that the true mean is not 17.  With 0.9 probability, this implies two boundaries; either the true mean is smaller than 17, with 0.05 probability spanning 0 to 0.05 and it is bigger than 17 with 0.05 probability spanning 0.95 to 1 so that the interior range represents 0.9 probability as required.

- **Critical Values:** The sample mean would have to be at least **$\pm$ `r abs(round(qt(0.05, 49), digits=3))` ** standard errors away from 17 to rule out a mean of 17 with 90% confidence.

- **Result:** Our sample mean is **2.14** standard errors away from 17.  The probability of something equally or **more extreme** is **0.037**.


### Case 2: A Lesser Alternative {data-commentary-width=400}


```{r, warning=FALSE, message=FALSE}
library(radiant)
result <- prob_tdist(df = 49, plb = 0.1)
p1a <- plot(result, type = "probs") + theme_minimal() + geom_label(aes(x=-1.299, y=0.25, label="t=-1.299")) +
  geom_label(aes(x=0, y=0.075, label="p=0.9"), size=2) +
  geom_label(aes(x=-3, y=0.075, label="p=0.1"), size=2) +
  labs(title="Critical Value") +
  xlim(-4,4)
result2 <- prob_tdist(df = 49, lb = -2.14)
p2a <- plot(result2) + theme_minimal() +
  geom_vline(aes(xintercept=-2.14), color="#099587") +
  geom_label(aes(x=-2.25, y=0.25, label="t=-2.14")) +
  geom_label(aes(x=0, y=0.075, label="p=0.981"), size=2) +
  geom_label(aes(x=-3, y=0.075, label="p=0.0187"), size=2) +
  guides(size="none") +
  labs(title="Result", x="t") + 
  xlim(-4,4)
library(patchwork)
p1a + p2a
```

***

``` 
Critical Values
P(X < -1.299) = 0.1
P(X > -1.299) = 0.9

Result
P(X < -2.14) = 0.019
P(X > -2.14) = 0.981
```

***

Knowing only the sample size, 50, is sufficient to determine what $t$ must be to reject a mean of 17 or greater in favor of the alternative that the true mean is less than 17.  With 0.9 probability, either the true mean is smaller than 17, with 0.1 probability or it is 17 or bigger with 0.9 probability as required.

- **Critical Values:** The sample mean would have to be at most **`r round(qt(0.1, 49), digits=3)` ** standard errors away from 17 to rule out a mean of 17 or greater with 90% confidence.

- **Result:** Our sample mean is **-2.14** standard errors away from 17.  The probability of something equal or **lesser** is **`r round(pt(-2.14, 49), digits=3)`**.


### Case 3: A Greater Alternative {data-commentary-width=400}


```{r, warning=FALSE, message=FALSE}
library(radiant)
result <- prob_tdist(df = 49, plb = 0.9)
p1a <- plot(result, type = "probs") + theme_minimal() + geom_label(aes(x=1.299, y=0.25, label="t=1.299")) +
  geom_label(aes(x=0, y=0.075, label="p=0.9"), size=2) +
  geom_label(aes(x=3, y=0.075, label="p=0.1"), size=2) +
  labs(title="Critical Value") +
  xlim(-4,4)
result2 <- prob_tdist(df = 49, lb = -2.14)
p2a <- plot(result2) + theme_minimal() +
  geom_vline(aes(xintercept=-2.14), color="#099587") +
  geom_label(aes(x=-2.25, y=0.25, label="t=-2.14")) +
  geom_label(aes(x=0, y=0.075, label="p=0.981"), size=2) +
  geom_label(aes(x=-3, y=0.075, label="p=0.0187"), size=2) +
  guides(size="none") +
  labs(title="Result", x="t") + 
  xlim(-4,4)
library(patchwork)
p1a + p2a
```

***

``` 
Critical Values
P(X < -1.299) = 0.1
P(X > -1.299) = 0.9

Result
P(X < -2.14) = 0.019
P(X > -2.14) = 0.981
```

***

Knowing only the sample size, 50, is sufficient to determine what $t$ must be to reject a mean of 17 or greater in favor of the alternative that the true mean is less than 17.  With 0.9 probability, either the true mean is smaller than 17, with 0.1 probability or it is 17 or bigger with 0.9 probability as required.

- **Critical Values:** The sample mean would have to be at most **`r round(qt(0.1, 49), digits=3)` ** standard errors away from 17 to rule out a mean of 17 or greater with 90% confidence.

- **Result:** Our sample mean is **-2.14** standard errors away from 17.  The probability of something equal or **greater** is **`r round(1-pt(-2.14, 49), digits=3)`**.

### `t.test` {data-commentary-width=600}

**Alternative: Two-sided**

```{r}
t.test(cars$speed, mu=17)
```

***

**Alternative: Less**

```{r}
t.test(cars$speed, mu=17, alternative="less")
```

**Alternative: Greater**

```{r}
t.test(cars$speed, mu=17, alternative = "greater")
```

