---
title: "OHLC Data and plotly"
bibliography: "bibliography.bib"
nocite: |
     @*
author: "Robert W. Walker"
format: 
   html:
     self-contained: true
     self-contained-math: true
     code-fold: true
     code-copy: hover
     code-block-border-left: true
     df-print: paged
     embed-resources: true
     fig-format: retina
     footnotes-hover: true
     html-math-method: katex
     mainfont: Playfair Display
     toc: true
     toc-title: Navigation
     theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, warning=FALSE)
```

# plotly for OHLC

OHLC data on Intel stock price.  The `tidyquant` packages makes the acquisition of stock market data easy by entering the ticker.  Let's try this for Intel, ticker **INTC**.

```{r}
library(tidyquant)
library(tidyverse)
library(magrittr)
# Use tidyquant to get the data
INTC <- tq_get("INTC")
# Slice off the most recent 120 days
# INTC.tail <- tail(INTC, 120)
INTC %<>% mutate(
    open = round(open, digits=2),
    close = round(close, digits=2),
    high = round(high, digits=2),
    low = round(low, digits=2),
    adjusted = round(adjusted, digits=2)
    )
```

Let's have a look at the data.

```{r}
library(DT)
datatable(INTC.tail)
```

## The Plot

There are a few charts specifically designed for OHLC data that are included in `plotly`.  Here I want to deploy a basic one with one modification.  I want daily increases in black and daily decreases in red.

```{r}
library(plotly)
# basic example of ohlc charts
# custom colors
i <- list(line = list(color = '#000000')) # black
d <- list(line = list(color = '#FF0000')) # red
# Create the figure
fig.2 <- INTC %>%
  plot_ly(x = ~date, type="ohlc",
          open = ~open, close = ~close,
          high = ~high, low = ~low,
          increasing = i, decreasing = d)
fig.2
```

# References

```{r}
knitr::write_bib(names(sessionInfo()$otherPkgs), file="bibliography.bib")
```

