---
title: "Probability Calculation in R"
author: "RWW"
format: 
   html:
     self-contained: true
     self-contained-math: true
     code-fold: true
     code-copy: hover
     code-block-border-left: true
     df-print: paged
     embed-resources: true
     fig-format: retina
     footnotes-hover: true
     html-math-method: katex
     mainfont: Playfair Display
     toc: true
     toc-title: Navigation
     theme: simplex
---

In an example problem in the slides, we have a drug test.  It is 99% accurate for Users and 95% accurate for non-Users.  We are told to assume that 10% of the population uses the drug in question.  Let's start by representing that latter fact.

```{r}
Marginal.Probabilities <- c(0.1,0.9)
names(Marginal.Probabilities) <- c("User","Non-User")
Marginal.Probabilities
```

Now I want to represent the conditional probabilities given above for the rest.

```{r}
Test.Conditional.Probabilities <- matrix(data=c(0.99,0.01,0.05,0.95), byrow=T, nrow=2, dimnames = list(c("User","Non-User"), c("Positive","Negative")))
Test.Conditional.Probabilities
```

From this, I can calculate the joint probabilities by multiplying a set of two conditional probabilities by the associated marginal probabilities.

```{r}
Joint <- Test.Conditional.Probabilities*Marginal.Probabilities
Joint
```

That gives me the table of joint probabilities that I sought.

Answering questions here, for example, what is the total proportion of positive tests?

```{r}
sum(Joint[,1])
```

How about the conditional probability of a User given a positive test?  That would be the joint probability of user and positive divided by the marginal probability of a positive test.  In language, it is the proportion of users (0.099) among all positives (0.144), the remainder are non-users that received positive test results.

```{r}
Joint[1,1]/sum(Joint[,1])
```

